# Multiple imputation  

 Multiple imputation is a general approach to the problem of missing data. It aims to allow for the uncertainty about the missing data by creating several different plausible imputed data sets and appropriately combining results obtained from each of them.    
   

 
 Multiple imputation using chained equations (MICE) were performed to generate 10 imputed datasets. For the imputation model, predictive mean matching was used for continuous data and logistic regression was used for binary data. 

<br/>

--------

<br/>

----
Using data **final_db**    
Outcome variable : **HTN**     
Follow-up period : **DATEDIFF**    
Exposure variable : **DM**     
Covariates : **Age, Sex, SES, Region, BMI, CCI, Comorbidities(Dyslipidemia, Ischemic heart disease)**   
----

<br/>

```{r include=FALSE}
library(mice)
library(dplyr)
```

```{r, eval=F}
## load library
library(mice)
library(dplyr)
```

```{r}
## load data
final_db <- read.csv('Data/final_db.csv', header=T)
```
     
----     

<br/> 
     
## The number of missing values
```{r}
na_count <- function(data){
  num.na <- colSums(is.na(data))
  per.na <- paste0(round(colSums(is.na(data))/nrow(data) *100,2),"%")
  
  return(data.frame(missing=paste0(num.na,"(",per.na,")"),row.names = names(num.na)))
}

na_count(final_db)
```

<br/>

* Use **mice()** function in **mice** package to deal with missing data.
  - m=10 refers to the number of imputed datasets. Five is the default value.
  - Extract imputed data sets using **compleate()** function  



## Imputation for missing values
```{r}
## Exclude subject ID, index date before imputation
dat_mice <- final_db %>% select(-RN_INDI, -INDEX_DT, -FU_DT) 
dat_imp <- mice(dat_mice, m=10, seed=1)
```

```{r}
## Create 10 imputed data
for (i in 1:dat_imp$m){
  z <- assign(paste0('dat_imp',i),complete(dat_imp,i))
  assign(paste0('dat_imp',i),cbind(z,final_db %>% select(RN_INDI))) 
}

## list of 10 imputed data 
dat_imp_list <- list(dat_imp1,dat_imp2,dat_imp3,dat_imp4,dat_imp5,dat_imp6,dat_imp7,dat_imp8,dat_imp9,dat_imp10)

```

```{r}
## Save multiple imputation result
save(dat_imp,file="Data/dat_imp.RData")
## Save list for imputed data
save(dat_imp_list,file="Data/dat_imp_list.RData")
```
