# Propensity Score Matching    
     

**Covariate balance check**    
    

Covariate balance is the degree to which the distribution of covariates is similar across levels of the treatment.    
SMD(Standardized Mean Difference) is the most widely used statistic for the assessment of balance after PSM.    

<br/>

SMD for continuous variables :   

$$SMD=\frac{\bar X_1-\bar X_2}{\sqrt{(S^2_1+S^2_2)/2}}$$

* $\bar X_1$ and $\bar X_2$ are sample mean for the treated and control groups.    
* $S^2_1$ and $S^2_2$ are sample variance for the treated and control groups.    

<br/>

SMD for binary variables :     

$$SMD=\frac{\hat p_1-\hat p_2}{\sqrt{[\hat p_1(1-\hat p_1)+\hat p_2(1-\hat p_2)]/2}}$$

* $\hat p_1$ and $\hat p_2$ are prevalence of binary variables in the treated and control groups.    

<br/>

If the SMD after matching is less than **0.1**, it is determined that the difference by the covariates between the two groups is negligible.     


<br/>

----
Using list **dat_imp_list**    
Outcome variable : **HTN**     
Follow-up period : **DATEDIFF**    
Exposure variable : **DM**     
Covariates : **Age, Sex, SES, Region, BMI, CCI, Comorbidities(Dyslipidemia, Ischemic heart disease)**   
----

<br/>

```{r include=FALSE}
library(MatchIt)
library(dplyr)
source("cobalt_3.9.0.R")
```

```{r, eval=F}
## load library
library(MatchIt)
library(dplyr)
source("cobalt_3.9.0.R")
```

```{r}
## load data
load("Data/dat_imp_list.RData")
final_com <- read.csv('Data/final_com.csv', header=T)
```

```{r}
## Formula
formula.mat <- formula(DM ~ AGE + SEX + SES + REGION + BMI + CCI + DYS + IHD)
```

<br/>

* Use **matchit()** function in **MatchIt** package to create treatment and control groups balanced on included covariates. 
  - method='nearest' : nearest neighbor matching on the propensity score
  - ratio=k : the number of controls matched to each treated unit for k:1 matching     
  - caliper : Units whose propensity score difference is larger than the caliper will not be paired, and some treated units may therefore not receive a match.


## Complete data version

**1:5 nearest matching _ caliper: 0.4**    

```{r}
## Optimal caliper
ps <- glm(formula.mat,data=final_com, family = 'binomial')
ps$pscore<- predict(ps, type='link')
0.2*sd(ps$pscore)
```

```{r}
set.seed(1)
mat <- matchit(formula.mat, method = 'nearest', data=final_com, ratio=5, caliper=0.4)
matdat <- match.data(mat) %>% select(-subclass) # 서버 R에서는 subclass 자동생성 안됨
    
# adding an matching index as a subclass to matdat and store as dat_mat
# as.numeric(tmp[,1:ratio]), rep(c(1:nrow(tmp)),ratio+1) 
tmp <- na.omit(mat$match.matrix)
matid <- data.frame(rowid=c(as.numeric(rownames(tmp)), as.numeric(tmp[,1:5])), subclass=rep(c(1:nrow(tmp)),6))
matid$RN_INDI <- final_com$RN_INDI[matid$row]
dat_mat <- matdat %>% left_join(matid %>% select(RN_INDI,subclass),by = 'RN_INDI') %>% filter(is.na(subclass)==F)
```

```{r}
## Save matching data
save(dat_mat,file="Data/dat_mat.RData")
```


## Balance check (Complete)

```{r}
bal.ch <- function(before_data, after_data, group){
  
  group<-deparse(substitute(group))
  
  # SMD before matching
  bal_check_un <- bal.tab.data.frame(before_data[covariates], treat=before_data[,group], binary="std", s.d.denom = "pooled")
  un <- abs(bal_check_un$Balance$Diff.Un)

  # SMD after matching
  bal_check_adj <- bal.tab.data.frame(after_data[covariates], treat=after_data[,group], binary="std", s.d.denom = "pooled")
  adj <- abs(bal_check_adj$Balance$Diff.Un)
  
  bal.res <- data.frame(un=round(un,3),adj=round(adj,3))
  rownames(bal.res) <- rownames(bal_check_un$Balance)
  return(bal.res)
}

covariates <- c("AGE","SEX","SES","REGION","BMI","CCI","DYS","IHD")
```

<br/>

**bal.ch(before_matching_data, after_matching_data, group variable)**    

```{r}
bal.ch(final_com, dat_mat, DM)
```


<br/>

## Missing data version

**1:3 nearest matching _ caliper: 0.3, 0.35, 0.4**

```{r}
## Optimal caliper
opt.clp <- c()
for (i in 1:length(dat_imp_list)){
  ps <- glm(formula.mat,data=dat_imp_list[[i]], family = 'binomial')
  ps$pscore<- predict(ps, type='link')
  opt.clp <- c(opt.clp, 0.2*sd(ps$pscore))
}
opt.clp; mean(opt.clp)
```


```{r}
caliper <- c(0.3,0.35,0.4)

for (i in 1:length(dat_imp_list)){
  for (j in caliper){
    set.seed(1)
    mat <- matchit(formula.mat, method = 'nearest', data=dat_imp_list[[i]], ratio=3, caliper=j)
    matdat <- match.data(mat) %>% select(-subclass) # 서버 R에서는 subclass 자동생성 안됨
    
    # adding an matching index as a subclass to matdat and store as dat_mati_j
    # as.numeric(tmp[,1:ratio]), rep(c(1:nrow(tmp)),ratio+1) 
    tmp <- na.omit(mat$match.matrix)
    matid <- data.frame(rowid=c(as.numeric(rownames(tmp)), as.numeric(tmp[,1:3])), subclass=rep(c(1:nrow(tmp)),4))
    matid$RN_INDI <- dat_imp_list[[i]]$RN_INDI[matid$row]
    assign(paste0('dat_mat',i,"_",j), matdat %>% left_join(matid %>% select(RN_INDI,subclass),by = 'RN_INDI') %>% filter(is.na(subclass)==F))
  }
}
```

```{r}
## list of 10 matched data 
dat_mat_list_0.3 <- list(dat_mat1_0.3,dat_mat2_0.3,dat_mat3_0.3,dat_mat4_0.3,dat_mat5_0.3,dat_mat6_0.3,dat_mat7_0.3,dat_mat8_0.3,dat_mat9_0.3,dat_mat10_0.3)
dat_mat_list_0.35 <- list(dat_mat1_0.35,dat_mat2_0.35,dat_mat3_0.35,dat_mat4_0.35,dat_mat5_0.35,dat_mat6_0.35,dat_mat7_0.35,dat_mat8_0.35,dat_mat9_0.35,dat_mat10_0.35)
dat_mat_list_0.4 <- list(dat_mat1_0.4,dat_mat2_0.4,dat_mat3_0.4,dat_mat4_0.4,dat_mat5_0.4,dat_mat6_0.4,dat_mat7_0.4,dat_mat8_0.4,dat_mat9_0.4,dat_mat10_0.4)

## Save list for matched data
save(dat_mat_list_0.3,file="Data/dat_mat_list_0.3.RData")
save(dat_mat_list_0.35,file="Data/dat_mat_list_0.35.RData")
save(dat_mat_list_0.4,file="Data/dat_mat_list_0.4.RData")
```


## Balance check (Complete)

```{r}

bal.ch <- function(dat_imp_list, dat_mat_list, group){
  
  group<-deparse(substitute(group))
  
  # SMD before matching
  bal_check_un <- dat_imp_list %>% lapply(function(x){
    bal.tab.data.frame(x[covariates], 
                       treat=x[,group], binary="std", s.d.denom = "pooled")})
  un <- sapply(bal_check_un, function(x) (abs(x$Balance$Diff.Un)))
  rownames(un) <- rownames(bal_check_un[[1]]$Balance)
  
  # SMD after matching
  bal_check_adj <- dat_mat_list %>% lapply(function(x){
    bal.tab.data.frame(x[covariates], 
                       treat=x[,group], binary="std", s.d.denom = "pooled")})
  adj <- sapply(bal_check_adj, function(x) (abs(x$Balance$Diff.Un)))
  rownames(adj) <- rownames(bal_check_adj[[1]]$Balance)
  
  bal.res <- list(un=apply(un, 1, summary), adj=apply(adj, 1, summary))
  return(data.frame(un=round(bal.res$un[6,],3),adj=round(bal.res$adj[6,],3)))
}

covariates <- c("AGE","SEX","SES","REGION","BMI","CCI","DYS","IHD")
```

<br/>

**bal.ch(before_matching_list, after_matching_list, group variable)**    

```{r}
bal.ch(dat_imp_list, dat_mat_list_0.3, DM)
bal.ch(dat_imp_list, dat_mat_list_0.35, DM)
bal.ch(dat_imp_list, dat_mat_list_0.4, DM)
```

