# Kaplan-Meier curve

## The Kaplan Meier estimator 

The Kaplan-Meier estimator is used to estimate the survival function. The visual representation of this function is usually called the Kaplan-Meier curve, and it shows what the probability of an event (for example, survival) is at a certain time interval.\
We compared two groups in this study. : DM(case) vs NonDM(control)\

<br/>

* Survival function :  

$$S(t)=P(T > t)$$

<br/>

* Kaplan-Meier estimator :   

$$\hat S(t)=\prod_{j:t_j\le t}\left(1-\frac{d_j}{n_j}\right)=\prod_{s\le t}\left(1-\frac{\bigtriangleup N(s)}{Y(s)}\right)$$


$d_j$ : the number of individuals who experience the event at $t_j$.\
$n_j$ : the number individuals at risk at $t_j$.\


<br/>

--------

<br/>

----
* Using data **final_com, dat_mat, dat_wt** (complete data version)
Using list **dat_mat_list, dat_imp_list** (missing data version)
Outcome variable : **HTN**     
Follow-up period : **DATEDIFF**    
Exposure variable : **DM**     
Covariates : **Age, Sex, SES, Region, BMI, CCI, Comorbidities(Dyslipidemia, Ischemic heart disease)**   
----

<br/>

```{r include=FALSE}
library(survival)
library(survminer)
library(ggsci)
library(RISCA)
```

```{r, eval=F}
## load library
library(survival)
library(survminer)
library(ggsci)
library(RISCA) # 서버에서는 library(IPWsurvival) 사용
```

```{r}
## load data
final_com <- read.csv('Data/final_com.csv', header=T)
load("Data/dat_mat.RData")
load("Data/dat_wt.RData")
load("Data/dat_mat_list_0.4.RData")
load("Data/dat_imp_list.RData")
```

```{r}
## Formula
formula.wt <- formula(DM ~ AGE + SEX + SES + REGION + BMI + CCI + DYS + IHD)
```

------

<br/>

## Complete data version

### Kaplan-Meier curve function

* Use **survfit(), ggsurvplot()** function in **survival, survminer** package to plot survival curve.    

<br/>

```{r}
gg_km <- function(data){
  p <- ggsurvplot(fit = fit_km,
                  title = "Kaplan-Meier Curve",
                  ggtheme = theme(axis.line = element_line(color="black"), 
                                  panel.background = element_blank(),
                                  plot.title = element_text(hjust = 0, size=16, face = "bold"),
                                  plot.margin = unit(c(5,5,5,5), "mm"),
                                  axis.title.y = element_text(margin=margin(r=13)),
                                  axis.title.x = element_text(margin=margin(t=10)),
                                  legend.text = element_text(size = 11.5, color="black"),
                                  legend.title = element_text(size=13.5)),
                  palette = c("#374E55FF", "#B24745FF"),
                  xlim = c(0,2500), 
                  ylim=c(0,1),
                  ### Censor Details 
                  censor = TRUE, # logical value. If TRUE, censors will be drawn
                  censor.shape="+", # Default value is "+", a sensible choice is "|".
                  censor.size = 4, 
                  ### Confidence Interval
                  conf.int = TRUE, # To Remove conf intervals use FALSE
                  ### Format Axes (changes x,y axis label)
                  xlab = "Time (Year)", 
                  ylab = "Survival probability",
                  font.x=c(15),
                  font.y=c(15),
                  font.xtickslab=c(11,"plain"), 
                  font.ytickslab=c(11,"plain"),
                  ### Format Legend
                  legend.title = "Diabetes Mellitus",
                  legend.labs = c("Control", "Case"), # Change the Strata Legend
                  legend = c(0.15,0.2), # c(0,0) corresponds to the "bottom left" and c(1,1) corresponds to the "top right" position
                  ### Risk Table 
                  risk.table = F, # To Remove risk table use FALSE
                  tables.height = 0.25, # Adjusts the height of the risk table
                  tables.col = "black",
                  tables.y.text = FALSE,
                  tables.y.text.col = TRUE,
                  risk.table.fontsize = 4.5,
                  tables.theme = theme_cleantable(font.y=12),
                  ### p-value details
                  pval = TRUE,
                  pval.size = 5,
                  pval.coord = c(2100,0.1)
  ) 
  
  p$plot <- p$plot + scale_x_continuous(breaks=seq(0,2600,365.25), labels=0:7) 
  return(p)
}
```

<br/>


### General version
```{r}
fit_km <- survfit(Surv(DATEDIFF, HTN==1)~ DM, data=final_com)
gg_km(fit_km)
```

### Matching version
```{r}
fit_km <- survfit(Surv(DATEDIFF, HTN==1)~ DM, data=dat_mat, weight = weights)
gg_km(fit_km)
```

### Weighting version
```{r}
fit_km <- survfit(Surv(DATEDIFF, HTN==1)~ DM, data=dat_wt, weight = weights)
gg_km(fit_km)
```


<br/>
<br/>

## Missing data version

### Matching version
```{r}
fit_km <- survfit(Surv(DATEDIFF, HTN==1)~ DM, data=dat_mat_list_0.4[[1]])
gg_km(fit_km)
```


### Weighting version

#### Weighted kaplan-meier estimates

* Use **ipw.survival()** function in **RISCA** package to estimate adjusted survival curves by weighting the individual contributions by the inverse of the probability to be in the group (IPW).   

```{r}
surv_res <- function(data, time, status, group, formula){
  p.score <- glm(formula=formula, data=data, family="binomial", x=TRUE, y=TRUE)$fitted.values
  wt <- (data[,group]=="1")/p.score + (data[,group]=="0")/(1-p.score)
  pt <- sum(data[,group]==1)/nrow(data)
  sw <- pt*(data[,group]==1) /p.score + (1-pt)*(data[,group]==0)/(1-p.score)
  res.akm <- ipw.survival(times=data[,time], failures=data[,status], variable=data[,group], weights=sw)
  return(res.akm)
}

res.akm <- dat_imp_list %>% lapply(function(x) surv_res(x,"DATEDIFF","HTN","DM", formula.wt)$table.surv)
risk <- do.call("cbind",res.akm %>% lapply(function(x) x$n.risk))
event <- do.call("cbind",res.akm %>% lapply(function(x) x$n.event))
surv <- do.call("cbind",res.akm %>% lapply(function(x) x$survival))
km_res <- data.frame(time=res.akm[[1]]$times, strata=res.akm[[1]]$variable,
                     n.risk=apply(risk,1,mean),
                     n.event=apply(event,1,mean),
                     surv=apply(surv,1,mean))
```

```{r}
head(km_res)
```


#### Kaplan-Meier curve by Using IPW

* Use **ggsurvplot_df()** function in **survminer** package  to plot survival curve from any data frame containing the summary of survival curves.   

```{r}
ggsurv_plot <- ggsurvplot_df(km_res,
                             title = "Kaplan-Meier Curve",
                             ggtheme = theme(axis.line = element_line(color="black"),
                                             panel.background = element_blank(),
                                             plot.title = element_text(hjust = 0, size=15, face = "bold"),
                                             plot.margin = unit(c(5,3,5,5), "mm"),
                                             axis.title.y = element_text(margin=margin(r=13)),
                                             axis.title.x = element_text(margin=margin(t=10)),
                                             legend.background =element_blank(),
                                             legend.text = element_text(size = 13, color="black"),
                                             legend.title = element_text( size=13.5),
                                             legend.spacing.x = unit(0.5, 'cm'),
                                             legend.spacing.y = unit(0.3, 'cm')),
                             palette = c("#374E55FF", "#B24745FF"),
                             xlim = c(0,2600),
                             ylim = c(0.5,1), 
                             size = 1.2,
                             ### Format Axes (changes x,y axis label)
                             xlab="Time (Years)", 
                             font.x = c(16), 
                             font.y = c(16),
                             font.tickslab = c(13),
                             ### Format Legend
                             legend.title = "Diabetes Mellitus:  ",
                             legend.labs = c("Control", "Case"), # Change the Strata Legend
                             legend = c(0.2,0.2) 
                             ) + scale_x_continuous(breaks = c(0,730.5,1461,2191.5,2922,3652.5), labels = c(0,2,4,6,8,10))
  
ggsurv_plot
```


